---
title: "casdas_2.0"
author: "Kate Pogue"
date: "2023-07-12"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Loading Packages
```{r}
library(ggplot2)
library(dplyr)
```


Loading Data
```{r}
NAB2 <- read_csv("data/NAB2_VER3.csv")
EPS <- read_excel("data/all_EPS.xlsx")
ASRT_main <- read_excel("data/ASRT_main.xlsx")
ASRT_other <- read_excel("data/ASRT_other.xlsx")
```


Merge some 4 and 9 data
```{r}
#ecoli
NAB2$`NAB4 Clarifier Effluent E. coli (LAB) (EC/100mL)`[is.na(NAB2$`NAB4 Clarifier Effluent E. coli (LAB) (EC/100mL)`)] <- 0
NAB2$`NAB9 Clarifier Effluent E. coli (LAB) (EC/100mL)`[is.na(NAB2$`NAB9 Clarifier Effluent E. coli (LAB) (EC/100mL)`)] <- 0

NAB2 <- mutate(NAB2, "NAB4 and 9 Clarifier Effluent E. coli (LAB) (EC/100mL)" = 
              (NAB2$`NAB4 Clarifier Effluent E. coli (LAB) (EC/100mL)` + NAB2$'NAB9 Clarifier Effluent E. coli (LAB) (EC/100mL)'))

NAB2$'NAB4 and 9 Clarifier Effluent E. coli (LAB) (EC/100mL)'[NAB2$'NAB4 and 9 Clarifier Effluent E. coli (LAB) (EC/100mL)' == 0] <- NA
NAB2 <- NAB2 %>% select(-'NAB4 Clarifier Effluent E. coli (LAB) (EC/100mL)', -'NAB9 Clarifier Effluent E. coli (LAB) (EC/100mL)')

# c-pass tknh

NAB2$`NAB4 C-Pass TKNH (LAB) (mg/L)`[is.na(NAB2$`NAB4 C-Pass TKNH (LAB) (mg/L)`)] <- 0
NAB2$`NAB9 C-Pass TKNH (LAB) (mg/L)`[is.na(NAB2$`NAB9 C-Pass TKNH (LAB) (mg/L)`)] <- 0

NAB2 <- mutate(NAB2, "NAB4 and 9 C-Pass TKNH (LAB) (mg/L)" = 
              (NAB2$`NAB4 C-Pass TKNH (LAB) (mg/L)` + NAB2$`NAB9 C-Pass TKNH (LAB) (mg/L)`))

NAB2$'NAB4 and 9 C-Pass TKNH (LAB) (mg/L)'[NAB2$'NAB4 and 9 C-Pass TKNH (LAB) (mg/L)' == 0] <- NA
NAB2 <- NAB2 %>% select(-'NAB4 C-Pass TKNH (LAB) (mg/L)', -'NAB9 C-Pass TKNH (LAB) (mg/L)')

# c-pass dsvi 30

NAB2$`NAB4 C-Pass dSVI 30 (mL/g)`[is.na(NAB2$`NAB4 C-Pass dSVI 30 (mL/g)`)] <- 0
NAB2$`NAB9 C-Pass dSVI 30 (mL/g)`[is.na(NAB2$`NAB9 C-Pass dSVI 30 (mL/g)`)] <- 0

NAB2 <- mutate(NAB2, "NAB4 and 9 C-Pass dSVI 30 (mL/g)" = 
              (NAB2$`NAB4 C-Pass dSVI 30 (mL/g)` + NAB2$`NAB9 C-Pass dSVI 30 (mL/g)`))

NAB2$'NAB4 and 9 C-Pass dSVI 30 (mL/g)'[NAB2$'NAB4 and 9 C-Pass dSVI 30 (mL/g)' == 0] <- NA
NAB2 <- NAB2 %>% select(-'NAB4 C-Pass dSVI 30 (mL/g)', -'NAB9 C-Pass dSVI 30 (mL/g)')

#c-pass dsvi 5

NAB2$`NAB4 C-Pass dSVI 5 (mL/g)`[is.na(NAB2$`NAB4 C-Pass dSVI 5 (mL/g)`)] <- 0
NAB2$`NAB9 C-Pass dSVI 5 (mL/g)`[is.na(NAB2$`NAB9 C-Pass dSVI 5 (mL/g)`)] <- 0

NAB2 <- mutate(NAB2, "NAB4 and 9 C-Pass dSVI 5 (mL/g)" = 
              (NAB2$`NAB4 C-Pass dSVI 5 (mL/g)` + NAB2$`NAB9 C-Pass dSVI 5 (mL/g)`))

NAB2$'NAB4 and 9 C-Pass dSVI 5 (mL/g)'[NAB2$'NAB4 and 9 C-Pass dSVI 5 (mL/g)' == 0] <- NA
NAB2 <- NAB2 %>% select(-'NAB4 C-Pass dSVI 5 (mL/g)', -'NAB9 C-Pass dSVI 5 (mL/g)')

# c-pass TSS


NAB2$`NAB9 C-Pass TSS (LAB) (mg/L)`[is.na(NAB2$`NAB9 C-Pass TSS (LAB) (mg/L)`)] <- 0
NAB2$`NAB4 C-Pass TSS (LAB) (mg/L)`[is.na(NAB2$`NAB4 C-Pass TSS (LAB) (mg/L)`)] <- 0

NAB2 <- mutate(NAB2, "NAB4 and 9 C-Pass TSS (LAB) (mg/L)" = 
              (NAB2$`NAB9 C-Pass TSS (LAB) (mg/L)` + NAB2$`NAB4 C-Pass TSS (LAB) (mg/L)`))

NAB2$'NAB4 and 9 C-Pass TSS (LAB) (mg/L)'[NAB2$'NAB4 and 9 C-Pass TSS (LAB) (mg/L)' == 0] <- NA
NAB2 <- NAB2 %>% select(-'NAB9 C-Pass TSS (LAB) (mg/L)', -'NAB4 C-Pass TSS (LAB) (mg/L)')

#ABXA NO5


NAB2$`AB4A NO5 (mg N/L)`[is.na(NAB2$`AB4A NO5 (mg N/L)`)] <- 0
NAB2$`AB9A NO5 (mg N/L)`[is.na(NAB2$`AB9A NO5 (mg N/L)`)] <- 0

NAB2 <- mutate(NAB2, "AB4 and 9A NO5 (mg N/L)" = 
              (NAB2$`AB4A NO5 (mg N/L)` + NAB2$`AB9A NO5 (mg N/L)`))

NAB2$'AB4 and 9A NO5 (mg N/L)'[NAB2$'AB4 and 9A NO5 (mg N/L)' == 0] <- NA
NAB2 <- NAB2 %>% select(-'AB4A NO5 (mg N/L)', -'AB9A NO5 (mg N/L)')

# ABXA NH3

NAB2$`AB4A NH3 (mg N/L)`[c(1:143, 297:435)] <- 0
NAB2$`AB9A NH3 (mg N/L)`[144:296] <- 0

NAB2 <- mutate(NAB2, "AB4 and 9A NH3 (mg N/L)" = 
              (NAB2$`AB4A NH3 (mg N/L)` + NAB2$`AB9A NH3 (mg N/L)`))

NAB2$'AB4 and 9A NH3 (mg N/L)'[NAB2$'AB4 and 9A NH3 (mg N/L)' == 0] <- NA
NAB2 <- NAB2 %>% select(-'AB4A NH3 (mg N/L)', -'AB9A NH3 (mg N/L)')

#ABXC NH3


NAB2$`AB4C NH3 (mg N/L)`[c(1:143, 298:435)] <- 0
NAB2$`AB9C NH3 (mg N/L)`[144:297] <- 0

NAB2 <- mutate(NAB2, "AB4 and 9C NH3 (mg N/L)" = 
              (NAB2$`AB4C NH3 (mg N/L)` + NAB2$`AB9C NH3 (mg N/L)`))

NAB2$'AB4 and 9C NH3 (mg N/L)'[NAB2$'AB4 and 9C NH3 (mg N/L)' == 0] <- NA
NAB2 <- NAB2 %>% select(-'AB4C NH3 (mg N/L)', -'AB9C NH3 (mg N/L)')

# ABX3 MLSS

NAB2$`AB4C MLSS (mg N/L)`[c(1:143, 297:435)] <- 0
NAB2$`AB9C MLSS (mg/L)`[143:296] <- 0

NAB2 <- mutate(NAB2, "AB4 and 9C MLSS (mg N/L)" = 
              (NAB2$`AB4C MLSS (mg N/L)` + NAB2$`AB9C MLSS (mg/L)`))


NAB2$'AB4 and 9C MLSS (mg N/L)'[NAB2$'AB4 and 9C MLSS (mg N/L)' == 0] <- NA
NAB2 <- NAB2 %>% select(-'AB4C MLSS (mg N/L)', -'AB9C MLSS (mg/L)')

#ABX INF FLOW


NAB2$`AB4 INF FLOW (MGD)`[c(1:143, 296:435)] <- 0
NAB2$`AB9 INF FLOW (MGD)`[144:295] <- 0

NAB2 <- mutate(NAB2, "AB4 and 9 INF FLOW (MGD)" = 
              (NAB2$`AB4 INF FLOW (MGD)` + NAB2$`AB9 INF FLOW (MGD)`))

NAB2$'AB4 and 9 INF FLOW (MGD)'[NAB2$'AB4 and 9 INF FLOW (MGD)' == 0] <- NA
NAB2 <- NAB2 %>% select(-'AB4 INF FLOW (MGD)', -'AB9 INF FLOW (MGD)')

#ABX MLR FLOW

NAB2$`AB4 MLR FLOW (MGD)`[297:435] <- 0
NAB2$`AB9 MLR FLOW (MGD)`[1:296] <- 0

NAB2 <- mutate(NAB2, "AB4 and 9 MLR FLOW (MGD)" = 
              (NAB2$`AB4 MLR FLOW (MGD)` + NAB2$`AB9 MLR FLOW (MGD)`))

NAB2$'AB4 and 9 MLR FLOW (MGD)'[NAB2$'AB4 and 9 MLR FLOW (MGD)' == 0] <- NA
NAB2 <- NAB2 %>% select(-'AB4 MLR FLOW (MGD)', -'AB9 MLR FLOW (MGD)')

#C-Pass VSS

NAB2$`NAB4 C-Pass VSS (LAB) (mg/L)`[is.na(NAB2$`NAB4 C-Pass VSS (LAB) (mg/L)`)] <- 0
NAB2$`NAB9 C-Pass VSS (LAB) (mg/L)`[is.na(NAB2$`NAB9 C-Pass VSS (LAB) (mg/L)`)] <- 0

NAB2 <- mutate(NAB2, "NAB4 and 9 C-Pass VSS (LAB) (mg/L)" = 
              (NAB2$`NAB4 C-Pass VSS (LAB) (mg/L)` + NAB2$`NAB9 C-Pass VSS (LAB) (mg/L)`))

NAB2$'NAB4 and 9 C-Pass VSS (LAB) (mg/L)'[NAB2$'NAB4 and 9 C-Pass VSS (LAB) (mg/L)' == 0] <- NA
NAB2 <- NAB2 %>% select(-'NAB4 C-Pass VSS (LAB) (mg/L)', -'NAB9 C-Pass VSS (LAB) (mg/L)')

#C-Pass total COD


NAB2$`NAB4 C-Pass Total COD (LAB) (mg/L)`[is.na(NAB2$`NAB4 C-Pass Total COD (LAB) (mg/L)`)] <- 0
NAB2$`NAB9 C-Pass Total COD (LAB) (mg/L)`[is.na(NAB2$`NAB9 C-Pass Total COD (LAB) (mg/L)`)] <- 0

NAB2 <- mutate(NAB2, "NAB4 and 9 C-Pass Total COD (LAB) (mg/L)" = 
              (NAB2$`NAB4 C-Pass Total COD (LAB) (mg/L)` + NAB2$`NAB9 C-Pass Total COD (LAB) (mg/L)`))

NAB2$'NAB4 and 9 C-Pass Total COD (LAB) (mg/L)'[NAB2$'NAB4 and 9 C-Pass Total COD (LAB) (mg/L)' == 0] <- NA
NAB2 <- NAB2 %>% select(-'NAB4 C-Pass Total COD (LAB) (mg/L)', -'NAB9 C-Pass Total COD (LAB) (mg/L)')


```


Create E. coli Reduction Variable and log transform everything
```{r}
## make LOG reduction variable
NAB2$log_pilot_reduction <- log(NAB2$`NSI E. coli (LAB) (EC/100mL)`, base = 10)- log(NAB2$`NAB2 Clarifier Effluent E. coli (LAB) (EC/100mL)`, base = 10)

NAB2$log_control_reduction <- log(NAB2$`NSI E. coli (LAB) (EC/100mL)`, base = 10)-log(NAB2$`NAB4 and 9 Clarifier Effluent E. coli (LAB) (EC/100mL)`, base = 10)


## normal effluent variables just need to be log transformed
NAB2$log_pilot_ecoli <- log(NAB2$`NAB2 Clarifier Effluent E. coli (LAB) (EC/100mL)`, base = 10)
NAB2$log_control_ecoli <- log(NAB2$`NAB4 and 9 Clarifier Effluent E. coli (LAB) (EC/100mL)`, base = 10)


#only select useful variables
NAB2 <- NAB2 %>% select(-c(`NSI E. coli (LAB) (EC/100mL)`, 
                           `NAB2 Clarifier Effluent E. coli (LAB) (EC/100mL)`,
                          `NAB4 and 9 Clarifier Effluent E. coli (LAB) (EC/100mL)`))

```


Select some Pilot factors
```{r}
# Convert "timestamp" column to a date vector
ASRT_main$timestamp <- as.Date(ASRT_main$timestamp, format = "%Y/%m/%d")
ASRT_other$timestamp <- as.Date(ASRT_other$timestamp, format = "%Y/%m/%d")
EPS <- EPS %>% rename(timestamp = 1)


#select pilot variables
pilot_eps <- EPS %>% select(timestamp, starts_with("Pilot"))
colnames(pilot_eps) <- c("timestamp", "pilot_total_cod_mgcod_per_gvss",
                  "pilot_total_pn_to_ps", "pilot_lb_cod_mgcod_per_gvss",
                  "pilot_lb_proteins_mg_per_gvss", "pilot_lb_carbs_mg_per_gvss",
                  "pilot_tb_cod_mgcod_per_gvss", "pilot_tb_proteins_mg_per_gvss",
                  "pilot_tb_carbs_mgcarb_per_gvss", "pilot_tb_pn_to_ps")

#get NAB2 factors
others <- NAB2 %>% select(timestamp,
                          "log_pilot_ecoli",
                          "log_pilot_reduction",
                          "Clarifier 2 SLR (ppd/sf)",                          
                          "Clarifier 2 SOR (gpd/sf)",                              
                          "NAB2 C-Pass Total COD (LAB) (mg/L)",
                          "NAB2 C-Pass TKNH (LAB) (mg/L)",
                          "NAB2 C-Pass TSS (LAB) (mg/L)",
                          "NAB2 C-Pass VSS (LAB) (mg/L)", 
                          "AB2A DO (mg/L)", 
                          "AB2B DO (mg/L)",
                          "AB2C DO (mg/L)",
                          "AB2A NO5 (mgN/L)", 
                          "AB2A NH3 (mg N/L)", 
                          "AB2C NH3 (mg N/L)" , 
                          "AB2C MLSS (mg/L)", 
                          "AB2 INF FLOW (MGD)", 
                          "AB2 MLR FLOW (MGD)", 
                          "AB2A AIRFLOW (SCFM)", 
                          "AB2B AIRFLOW (SCFM)", 
                          "AB2C AIRFLOW (SCFM)", 
                          "SC2 BLANKET (ft)", 
                          "NAB2 C-Pass dSVI 30 (mL/g)", 
                          "NAB2 C-Pass dSVI 5 (mL/g)"  
                          ,"NAB2 Clarifier Effluent Total COD (Low Level) (LAB) (mg/L)"  
,"NAB2 Clarifier Effluent NH3 (LAB) (mg/L)"                  
,"NAB2 Clarifier Effluent NO2 (LAB) (mg/L)"                  
,"NAB2 Clarifier Effluent NO5 (LAB) (mg/L)"                  
,"NAB2 Clarifier Effluent OP (LAB) (mg/L)"                             
,"NAB2 Clarifier Effluent PWS (LAB) (mg/L)"                  
,"NAB2 Clarifier Effluent PW (LAB) (mg/L)"                   
,"NAB2 Clarifier Effluent TKNH (LAB) (mg/L)"                 
,"NAB2 Clarifier Effluent TSS (LAB) (mg/L)"                  
,"NAB2 Clarifier Effluent VSS (LAB) (mg/L)"
                          )

colnames(others) <- c(
  "timestamp",
  "log_ecoli",
  "log_reduction",
  "CLARIFIER_2_SLR",
  "CLARIFIER_2_SOR",
  "NAB2_CPASS_TOTAL_COD",
  "NAB2_CPASS_TKNH",
  "NAB2_CPASS_TSS",
  "NAB2_CPASS_VSS",
  "AB2A_DO",
  "AB2B_DO",
  "AB2C_DO",
  "AB2A_NO5",
  "AB2A_NH3",
  "AB2C_NH3",
  "AB2C_MLSS",
  "AB2_INF_FLOW",
  "AB2_MLR_FLOW",
  "AB2A_AIRFLOW",
  "AB2B_AIRFLOW",
  "AB2C_AIRFLOW",
  "SC2_BLANKET",
  "NAB2_CPASS_dSVI30",
  "NAB2_CPASS_dSVI5",
  "NAB2_CLARIFIER_EFFLUENT_TOTAL_COD",
  "NAB2_CLARIFIER_EFFLUENT_NH3",
  "NAB2_CLARIFIER_EFFLUENT_NO2",
  "NAB2_CLARIFIER_EFFLUENT_NO5",
  "NAB2_CLARIFIER_EFFLUENT_OP",
  "NAB2_CLARIFIER_EFFLUENT_PWS",
  "NAB2_CLARIFIER_EFFLUENT_PW",
  "NAB2_CLARIFIER_EFFLUENT_TKNH",
  "NAB2_CLARIFIER_EFFLUENT_TSS",
  "NAB2_CLARIFIER_EFFLUENT_VSS"
)


#get ASRT-other variables for pilot
ASRT_other_pilot <- ASRT_other %>% select(timestamp, starts_with("Pilot"))
colnames(ASRT_other_pilot) <- c("timestamp",
                                "Pilot_ASRT_d",
                                "Pilot_Floc_SRT_d",
                                "Pilot_Granule_SRT_d",
                                "Pilot_AGS_Weighted_SRT_d",
                                "Pilot_Floc_percent",
                                "Pilot_Granule_percent")

#Join
pilot_eps <- left_join(pilot_eps, others, by = "timestamp")
pilot_eps <- left_join(pilot_eps, ASRT_main[,-3], by = "timestamp")
pilot_eps <- left_join(pilot_eps, ASRT_other_pilot, by = "timestamp")

```

Select Control Factors
```{r}
#select control variables
control_eps <- EPS %>% select(timestamp, starts_with("Control"))
colnames(control_eps) <- c("timestamp", "control_total_cod_mgcod_per_gvss",
                  "control_total_pn_to_ps", "control_lb_cod_mgcod_per_gvss",
                  "control_lb_proteins_mg_per_gvss", "control_lb_carbs_mg_per_gvss",
                  "control_tb_cod_mgcod_per_gvss", "control_tb_proteins_mg_per_gvss",
                  "control_tb_carbs_mgcarb_per_gvss", "control_tb_pn_to_ps")

#get NAB2 factors
others <- NAB2 %>% select(timestamp,
"log_control_ecoli",
"log_control_reduction",
"Clarifier 9 SLR (ppd/sf)"                                  
,"Clarifier 9 SOR (gpd/sf)" 
,"AB9A DO (mg/L)"      
,"AB9B DO (mg/L)"                                            
,"AB9C DO (mg/L)" 
,"AB9A AIRFLOW (SCFM)"   
,"AB9B AIRFLOW (SCFM)"                                       
,"AB9C AIRFLOW (SCFM)"   
,"SC9 BLANKET (ft)"   
,"NNE Total COD (Low Level) (LAB) (mg/L)"                     
,"NNE NH3A (LAB) (mg/L)"                                     
,"NNE NO2 (LAB) (mg/L)"                                       
,"NNE NO5 (LAB) (mg/L)"                                      
,"NNE OP (LAB) (mg/L)"                                        
,"NNE PWS (LAB) (mg/L)"                                      
,"NNE PW (LAB) (mg/L)"                                       
,"NNE TKNH (LAB) (mg/L)"                                     
,"NNE TSS (LAB) (mg/L)"                                       
,"NNE VSS (LAB) (mg/L)"
,"NAB4 and 9 C-Pass Total COD (LAB) (mg/L)"
,"NAB4 and 9 C-Pass VSS (LAB) (mg/L)"
,"NAB4 and 9 C-Pass TKNH (LAB) (mg/L)"
,"NAB4 and 9 C-Pass dSVI 30 (mL/g)"                                
,"NAB4 and 9 C-Pass dSVI 5 (mL/g)"
,"NAB4 and 9 C-Pass TSS (LAB) (mg/L)"
,"AB4 and 9A NO5 (mg N/L)"
,"AB4 and 9A NH3 (mg N/L)" 
,"AB4 and 9C NH3 (mg N/L)" 
,"AB4 and 9C MLSS (mg N/L)" 
,"AB4 and 9 INF FLOW (MGD)"  
,"AB4 and 9 MLR FLOW (MGD)"
                          )

colnames(others) <- new_names_vector <- c(
  "timestamp",
  "log_ecoli",
  "log_reduction",
  "CLARIFIER_9_SLR",
  "CLARIFIER_9_SOR",
  "AB9A_DO",
  "AB9B_DO",
  "AB9C_DO",
  "AB9A_AIRFLOW",
  "AB9B_AIRFLOW",
  "AB9C_AIRFLOW",
  "SC9_BLANKET",
  "NNE_TOTAL_COD",
  "NNE_NH3A",
  "NNE_NO2",
  "NNE_NO5",
  "NNE_OP",
  "NNE_PWS",
  "NNE_PW",
  "NNE_TKNH",
  "NNE_TSS",
  "NNE_VSS",
  "NAB4_and_9_CPASS_TOTAL_COD",
  "NAB4_and_9_CPASS_VSS",
  "NAB4_and_9_CPASS_TKNH",
  "NAB4_and_9_CPASS_dSVI30",
  "NAB4_and_9_CPASS_dSVI5",
  "NAB4_and_9_CPASS_TSS",
  "AB4_and_9A_NO5",
  "AB4_and_9A_NH3",
  "AB4_and_9C_NH3",
  "AB4_and_9C_MLSS",
  "AB4_and_9_INF_FLOW",
  "AB4_and_9_MLR_FLOW"
)

# Reorder to match correspongind pilot variables
others <- others %>%
  select(
    timestamp,
    log_ecoli,
    log_reduction,
    CLARIFIER_9_SLR,
    CLARIFIER_9_SOR,
    NAB4_and_9_CPASS_TOTAL_COD,
    NAB4_and_9_CPASS_TKNH,
    NAB4_and_9_CPASS_TSS,
    NAB4_and_9_CPASS_VSS,
    AB9A_DO,
    AB9B_DO,
    AB9C_DO,
    AB4_and_9A_NO5,
    AB4_and_9A_NH3,
    AB4_and_9C_NH3,
    AB4_and_9C_MLSS,
    AB4_and_9_INF_FLOW,
    AB4_and_9_MLR_FLOW,
    AB9A_AIRFLOW,
    AB9B_AIRFLOW,
    AB9C_AIRFLOW,
    SC9_BLANKET,
    NAB4_and_9_CPASS_dSVI30,
    NAB4_and_9_CPASS_dSVI5,
    NNE_TOTAL_COD,
    NNE_NH3A,
    NNE_NO2,
    NNE_NO5,
    NNE_OP,
    NNE_PWS,
    NNE_PW,
    NNE_TKNH,
    NNE_TSS,
    NNE_VSS
  )

#get other ASRT variables for pilot
ASRT_other_control <- ASRT_other %>% select(timestamp, starts_with("Control"))
colnames(ASRT_other_control) <- c("timestamp",
                                "Control_ASRT_d",
                                "Control_Floc_SRT_d",
                                "Control_Granule_SRT_d",
                                "Control_AGS_Weighted_SRT_d",
                                "Control_Floc_percent",
                                "Control_Granule_percent")


#FILL method to fill in missing EPS and ASRT values
control_eps <- left_join(control_eps, others, by = "timestamp")
control_eps <- left_join(control_eps, ASRT_main[,-2], by = "timestamp")
control_eps <- left_join(control_eps, ASRT_other_control, by = "timestamp")


```

Combine dataframes with factorial variable
```{r}
# Mutate 'system' variable to be a factor with all 2s for pilot and all 9s for control
pilot_eps<- pilot_eps %>%
  mutate(system = factor(rep(2, nrow(pilot_eps))))

control_eps <- control_eps %>%
  mutate(system = factor(rep(9, nrow(control_eps))))


new_column_names <- c(
"timestamp"                              
,"total_cod_mgcod_per_gvss"        
,"total_pn_to_ps"                   
,"lb_cod_mgcod_per_gvss"           
,"lb_proteins_mg_per_gvss"          
,"lb_carbs_mg_per_gvss"            
,"tb_cod_mgcod_per_gvss"            
,"tb_proteins_mg_per_gvss"         
,"tb_carbs_mgcarb_per_gvss"         
,"tb_pn_to_ps"      
,"log_ecoli"
,"log_reduction"
,"CLARIFIER_X_SLR"                        
,"CLARIFIER_X_SOR"                       
,"NABX_CPASS_TOTAL_COD"                   
,"NABX_CPASS_TKNH"                       
,"NABX_CPASS_TSS"                         
,"NABX_CPASS_VSS"                        
,"ABXA_DO"                                
,"ABXB_DO"                               
,"ABXC_DO"                                
,"ABXA_NO5"                              
,"ABXA_NH3"                               
,"ABXC_NH3"                              
,"ABXC_MLSS"                              
,"ABX_INF_FLOW"                          
,"ABX_MLR_FLOW"                           
,"ABXA_AIRFLOW"                          
,"ABXB_AIRFLOW"                           
,"ABXC_AIRFLOW"                          
,"SCX_BLANKET"                            
,"NABX_CPASS_dSVI30"                     
,"NABX_CPASS_dSVI5"                       
,"NABX_CLARIFIER_EFFLUENT_TOTAL_COD"     
,"NABX_CLARIFIER_EFFLUENT_NH3"            
,"NABX_CLARIFIER_EFFLUENT_NO2"           
,"NABX_CLARIFIER_EFFLUENT_NO5"            
,"NABX_CLARIFIER_EFFLUENT_OP"            
,"NABX_CLARIFIER_EFFLUENT_PWS"            
,"NABX_CLARIFIER_EFFLUENT_PW"            
,"NABX_CLARIFIER_EFFLUENT_TKNH"           
,"NABX_CLARIFIER_EFFLUENT_TSS"           
,"NABX_CLARIFIER_EFFLUENT_VSS"            
,"NAB2_ASRT"                       
,"ASRT_d"                           
,"Floc_SRT_d"                      
,"Granule_SRT_d"                    
,"AGS_Weighted_SRT_d"              
,"Floc_percent"                     
,"Granule_percent"  
,"system"
                      )


pilot_eps <- setNames(pilot_eps, new_column_names)
control_eps <- setNames(control_eps, new_column_names)

# Stack the dataframes vertically and change column names
all_casdas <- rbind(pilot_eps, control_eps)


# Turn zeroing out to NA values
all_casdas$'ABXA_NO5'[all_casdas$'ABXA_NO5' == 0] <- NA
all_casdas$'ABXA_NH3'[all_casdas$'ABXA_NH3' == 0] <- NA
all_casdas$'ABXC_MLSS'[all_casdas$'ABXC_MLSS' == 0] <- NA
all_casdas$'ABX_INF_FLOW'[all_casdas$'ABX_INF_FLOW' == 0] <- NA
all_casdas$'ABX_MLR_FLOW'[all_casdas$'ABX_MLR_FLOW' == 0] <- NA

```


Make Distributions
```{r}

# Get the variable names from columns 2-50
variable_names <- names(all_casdas)[2:50]

# Perform comparison for each variable
for (variable in variable_names) {
  # Filter data for system 2 and system 9
  system_2 <- all_casdas[all_casdas$system == 2, variable]
  system_9 <- all_casdas[all_casdas$system == 9, variable]

  # Create density plots
  plot <- ggplot(all_casdas, aes(x = .data[[variable]], fill = factor(system))) +
    geom_density(alpha = 0.5) +
    labs(x = variable, y = "Density", title = paste("Density Curves of", variable)) +
    scale_fill_discrete(name = "System")

  # Print the plot
  print(plot)
}


```

Make seperate plot for ecoli
```{r}

# Convert 'system' to a factor variable
all_casdas$system <- factor(all_casdas$system, levels = c(2, 9), labels = c("Pilot", "Control"))

# Create a boxplot using ggplot2 with updated labels and colors
ggplot(all_casdas, aes(x = system, y = log_ecoli, fill = system)) +
  geom_boxplot() +
  labs(x = "System", y = "Log Transformed E. coli Concentration (EC/100mL)",
       title = "Effluent E. coli Concentration by System") +
  scale_fill_manual(values = c("Pilot" = "#2596be", "Control" = "#eeb26f"))


```





Make t-tests in table
```{r}
# Create an empty dataframe to store the results
t_tests <- data.frame(
  Variable = character(),
  T_Statistic = numeric(),
  P_Value = numeric(),
  stringsAsFactors = FALSE
)

# Get the variable names from columns 2-49
variable_names <- names(all_casdas)[2:50]

# Perform comparison for each variable
for (variable in variable_names) {
  # Filter data for system 2 and system 9
  system_2 <- all_casdas[all_casdas$system == 2, variable]
  system_9 <- all_casdas[all_casdas$system == 9, variable]

  # Perform two-sample t-test
  t_test_result <- t.test(system_2, system_9)
  
  # Extract relevant information from the t-test result
  variable_name <- variable
  t_statistic <- t_test_result$statistic
  p_value <- t_test_result$p.value
  
  # Create a new row in the results dataframe
  new_row <- data.frame(
    Variable = variable_name,
    T_Statistic = t_statistic,
    P_Value = p_value,
    stringsAsFactors = FALSE
  )
  
  # Append the new row to the results dataframe
  t_tests <- rbind(t_tests, new_row)
}

# Print the results dataframe
print(t_tests)


```


Confidence intervals
```{r}
# Create an empty dataframe to store the results
confidence_intervals <- data.frame(
  Variable = character(),
  Percentile_0.025_System_9 = numeric(),
  Percentile_0.975_System_9 = numeric(),
  Percentile_0.025_System_2 = numeric(),
  Percentile_0.975_System_2 = numeric(),
  stringsAsFactors = FALSE
)

# Get the variable names from columns 2-50
variable_names <- names(all_casdas)[2:50]

# Perform calculation for each variable
for (variable in variable_names) {
  # Filter data for system 2 and system 9
  system_2 <- all_casdas[all_casdas$system == 2, variable]
  system_9 <- all_casdas[all_casdas$system == 9, variable]

  # Calculate percentiles for system 9
  percentile_0.025_system_9 <- quantile(system_9, 0.025, na.rm = TRUE)
  percentile_0.975_system_9 <- quantile(system_9, 0.975, na.rm = TRUE)

  # Calculate percentiles for system 2
  percentile_0.025_system_2 <- quantile(system_2, 0.025, na.rm = TRUE)
  percentile_0.975_system_2 <- quantile(system_2, 0.975, na.rm = TRUE)

  # Create a new row in the results dataframe
  new_row <- data.frame(
    Variable = variable,
    Percentile_0.025_System_9 = percentile_0.025_system_9,
    Percentile_0.975_System_9 = percentile_0.975_system_9,
    Percentile_0.025_System_2 = percentile_0.025_system_2,
    Percentile_0.975_System_2 = percentile_0.975_system_2,
    stringsAsFactors = FALSE
  )

  # Append the new row to the results dataframe
  confidence_intervals <- rbind(confidence_intervals, new_row)
}

# Print the results dataframe
print(confidence_intervals)

```


Make Boxplot to compare lasso selections
```{r}
# Subset the data for control and pilot
pilot <- all_casdas[all_casdas$system == 2, ]
control <- all_casdas[all_casdas$system == 9, ]


# Pn to ps
variable_name <- "total_pn_to_ps"

# Create a new plot
plot1 <- ggplot() +
  # Boxplots
  geom_boxplot(data = rbind(pilot, control), aes(x = factor(system), y = .data[[variable_name]])) +
  # Axis labels and plot title
  labs(x = "System", y = "Total Protein to Carb Ratio (mg/mg)") +
  ggtitle(paste("Comparison of", "Total Protein to Carb Ratio", "between System 2 and System 9")) +
  # Change x-axis labels
  scale_x_discrete(labels = c("Pilot", "Control")) +
  # Modify the theme
  theme(panel.background = element_rect(fill = "#FF9A7B"))

print(plot1)

#ABA airflow
variable_name <- "ABXA_AIRFLOW"

# Create a new plot
plot2 <- ggplot() +
  # Boxplots
  geom_boxplot(data = rbind(pilot, control), aes(x = factor(system), y = .data[[variable_name]])) +
  # Axis labels and plot title
  labs(x = "System", y = "Aeration Basin A Airflow (SCFM)") +
  ggtitle(paste("Comparison of", "Aeration Basin A Airflow", "between System 2 and System 9")) +
  # Change x-axis labels
  scale_x_discrete(labels = c("Pilot", "Control")) +
  # Modify the theme
  theme(panel.background = element_rect(fill = "#FF9A7B"))

#ABA airflow
variable_name <- "ABXC_AIRFLOW"

# Create a new plot
plot3 <- ggplot() +
  # Boxplots
  geom_boxplot(data = rbind(pilot, control), aes(x = factor(system), y = .data[[variable_name]])) +
  # Axis labels and plot title
  labs(x = "System", y = "Aeration Basin C Airflow (SCFM)") +
  ggtitle(paste("Comparison of", "Aeration Basin C Airflow", "between System 2 and System 9")) +
  # Change x-axis labels
  scale_x_discrete(labels = c("Pilot", "Control")) +
  # Modify the theme
  theme(panel.background = element_rect(fill = "#FF9A7B"))


#NABX_CLARIFIER_EFFLUENT_NO5
variable_name <- "NABX_CLARIFIER_EFFLUENT_NO5"

# Create a new plot
plot4 <- ggplot() +
  # Boxplots
  geom_boxplot(data = rbind(pilot, control), aes(x = factor(system), y = .data[[variable_name]])) +
  # Axis labels and plot title
  labs(x = "System", y = "Clarifier Effluent NO5 (mg/L)") +
  ggtitle(paste("Comparison of", "Clarifier Effluent NO5", "between System 2 and System 9")) +
  # Change x-axis labels
  scale_x_discrete(labels = c("Pilot", "Control")) +
  # Modify the theme
  theme(panel.background = element_rect(fill = "#FF9A7B"))


#Floc_percent
variable_name <- "Floc_percent"

# Create a new plot
plot5 <- ggplot() +
  # Boxplots
  geom_boxplot(data = rbind(pilot, control), aes(x = factor(system), y = .data[[variable_name]])) +
  # Axis labels and plot title
  labs(x = "System", y = "Floc Mass Fraction (%)") +
  ggtitle(paste("Comparison of", "Floc Mass Fraction", "between System 2 and System 9")) +
  # Change x-axis labels
  scale_x_discrete(labels = c("Pilot", "Control")) +
  # Modify the theme
  theme(panel.background = element_rect(fill = "#D1FFC9"))

#Granule_percent
variable_name <- "Granule_percent"

# Create a new plot
plot6 <- ggplot() +
  # Boxplots
  geom_boxplot(data = rbind(pilot, control), aes(x = factor(system), y = .data[[variable_name]])) +
  # Axis labels and plot title
  labs(x = "System", y = "Granule Mass Fraction (%)") +
  ggtitle(paste("Comparison of", "Granule Mass Fraction", "between System 2 and System 9")) +
  # Change x-axis labels
  scale_x_discrete(labels = c("Pilot", "Control")) +
  # Modify the theme
  theme(panel.background = element_rect(fill = "#FF9A7B"))



print(plot1)
print(plot2)
print(plot3)
print(plot4)
print(plot5)
print(plot6)

```


Make Boxplots for confidence interval 
